On the application of microwave acoustic resonant viral inactivation \thanks{I would be delighted to include any criticisms or comments anyone may have; preferably leave them on the GitHub issues page, or email therobotist@gmail.com, @0xDBFB7 on Twitter, or irc.0xDBFB7.com:6667 \#covid.}

### 2020-06-10 documents/paper.tex
Some more documentation

\title{On the application of microwave acoustic resonant viral inactivation
\thanks{I would be delighted to include any criticisms or comments anyone may have; preferably leave them on the GitHub issues page, or email therobotist@gmail.com, @0xDBFB7 on Twitter, or irc.0xDBFB7.com:6667 \#covid.}}
\date{May 2020}
\author{Based primarily on exceptional work by }

@@

Testing was only performed with a surrogate bacteriophage. All experiments must be repeated with SARS-NCoV-2.

@@

Contrary to Far UV, where ozone production.

@@

Because this is an ostensibly novel and niche mechanism, it may behoove us to breifly review the biological basis for the safety limits set by standards organizations. \footnote{The idea of biological microwave resonances appears to have originated in [Frohlich 1980].}

@@

There are some 20,000 papers on the topic of microwave safety, with solid in vitro, in vivo and epidemiological evidence of safety in the 2.4 GHz and 5.8 GHz bands.

However, "There are limited experimental human data upon which to set limits on exposures above 6 GHz" [Chan 2019]. The frequency ranges above 6 GHz appear to be set primarily by assuming purely thermal mechanisms and introducing a substantial safety factor of 20 or 50x.

The IEEE standard which the flagship paper [Yang 2015] references [IEEE C95.1–2005] was recently overhauled [IEEE C95.1-2019] with new metrics. The

@@

[Adair 2002] demonstrate theoretically that acoustic resonances are not possible in biological solvents. All resonance modes are strongly overdamped by the surrounding solvent, and no amplitude amplification can take place.

This analysis would seem to disagree with the viral resonance apparently conclusively demonstrated by [Yang 2015]; the PCR result seems especially inarguable. We cannot explain this discrepancy.

[Liu 2009] cite [Edwards, Swicord 1984], saying: "It was proposed that a hydration layer surrounding DNA molecules could lower the viscoelastic transition
frequency, raise the quality factor of confined acoustic vibrations, and result in a microwave resonant absorption". But [Adair 2002]'s [Foster 1987] seem to quite conclusively put the kibosh on that idea, demonstrating no resonances and showing previous results to be an artifact of the measurement equipment.

However, [Sagripanti 1986] find DNA damage at $8.5 mW/g$ at 2.55 GHz, with only 0.8C heating of the solvent.

@@

9.4 GHz pulsed / in vivo, mouse, $0.1-10 \text{mW}/\text{m}^2$ over the whole body for 2 weeks.

@@

More modern research does not tend to agree with these concerns. As cited by [ICNIRP 2020]'s excellent literature review, [Vijayalaxmi 2006] expose lymphocyte blood cells to pulsed 8.2 GHz radiation for 2 hours at an average power of $10 \text{mW/cm}^2$. They find no change in any of the parameters measured, including various chromosomal parameters. They further discuss previous results.

We must defer to health experts for interpretation of these data, and will not attempt to draw a conclusion.

It would be ideal to have an expert review our claims before publication. However, we tenatively make the following claims:

@@

Evidence of such microwave absorption resonances has been reported ([Edwards et al., 1984] [Grundler and Keilman, 1983], but these results have not been replicated ([Gabriel et al., 1987] [Foster et al., 1987]

> We extend analyses that suggest that the damping of the vibratory motion by biological fluids severely restrict such possibilities[].

[Foster (1987)] find that the DNA resonance found by [Edwards (1987)] was well below the experimental error, given the latter's setup. The Foster paper is of excellent quality.

 What is that in $mw/cm^2$?

Via the :

> 
> Similarly, as there is no evidence that continuous (e.g., sinusoidal) and discontinuous (e.g., pulsed) EMFs result in different biological effects (Kowalczuk et al. 2010; Juutilainen et al. 2011), no theoretical distinction has been made between these types of exposure (all exposures have been considered empirically in terms of whether they adversely affect health)

Juutilainen references a wealth of empirical data. Vijayalaxmi (2006) is particularly helpful:

Chemeris discusses 8.8 GHz  pulses, and finds no effect:

> We found no data on genotoxic effects of HPPP EMF at peak SAR exceeding hundreds of megawatts per kilogram and at average SAR, which does not induce dramatic overheating of the exposed objects. At such powerful EMF, peak electric field intensity in the exposed biological objects is close to maximally achievable, i.e. near breakdown electric field intensity. It would be reasonable to expect specific genotoxic effects of RF EMF under such extreme conditions.

@@

[Hand 1982] $E_{mag}=1/e$ (13.5\% power density) skin\footnote{electromagnetic skin, not tissue skin}\footnote{well, both, I suppose.} depth of 10 GHz microwave is approximately:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
F=10 GHz          & Dry tissue & Wet tissue \\ \hline
Penetration depth & 30 mm      & $\sim$5 mm \\ \hline
\end{tabular}
\end{center}

@@

However, destruction of lung tissue appears to be the 	[Yuan 2019]

@@

For ease of design and simulation, a device with Touchstone S-parameters and SPICE files is greatly preferable.
%
\paragraph{\textbf{The feedback loop oscillator}}\

@@

With 0.79 mm FR4 substrate and 0.2 mm (8 mil) wide traces, the maximum impedance achievable was about 115 ohms, which did not appear to be sufficient as an RF choke.

If suitably high-impedance traces are not available, a common technique is to use a quarter-wavelength line (approximately 6 mm long with the above parameters) terminated with a stub to produce a virtual short or open circuit [Seo 2007].

@@

@@

Professor Shahramian's excellent video blog {\it The Signal Path} inspired this work.

@@

They say: "Based on the IEEE Microwave Safety Standard, the spatial averaged value of the power density in air in open public space shall not exceed the equivalent power density of $100(f/3)^{1/5} \text{ W/m}^2$ at frequencies between 3 and 96GHz ($f$ is in GHz). This corresponds to $115 \text{ W}/\text{m}^2$ at 6 GHz". The 115 W @ 6 GHz value corresponds to this equation with a coefficient of 100.

@@

We don't think this changes the substance of [Yang 2015]'s paper, if it is an error; it just means the mechanism is more useful for healthcare workers who acknowledge the risks.

@@

Simulating the oscillator with an AC sim using the manufacturer's S-parameters and QUCS' microstrip approximations [qucs/optimize\_filter\_1] appeared to yield good agreement with experiment. The peaks in the feedback voltage simulation corresponded approximately with peaks in the observed spectrum. [figures from LO prototype N].

@@

A filter design with apparently ideal phase and frequency properties was designed using this process, but again poor agreement was found with experiment.

@@

Because we lacked a spectrum analyzer that could monitor the higher poles of the filter until the bootstrap LO was designed.

@@

QUCS' microstrip models are not yet compatible with transient simulations; and some improved filter designs required simulating coupling between more than two microstrips, which QUCS did not yet support natively. Coupling can be emulated by introducing discrete coupling capacitors; but

@@

It's possible to link SPICE and FDTD in two main ways

@@

@@

It may be helpful to treat simulations very similarly to IRL experiments.

This is obvious to all compentent, but when rapidly iteratively testing with simulations, it may be helpful to automatically save a package with images of all the components (schematics, graphs) of each distinct test. Including a picture with a webcam of the assembled board is also helpful.

Version control alone isn't quite enough. Just having a simulation setup file somewhere in the commit history isn't "discoverable" - that is, you should be able to see what the input and output was without re-running the simulation.

@@

Sacred, recipy, Sumatra

@@

For instance, this was done with the PDB reader.

@@

@@

We present this only as a cautionary tale in the hopes that someday I will listen, rather than talk.

>>>>

\title{On the application of microwave acoustic resonant viral inactivation
\thanks{I would be delighted to include any criticisms or comments anyone may have, both on substance and comprehensibility; preferably leave them on the GitHub issues page, or email therobotist@gmail.com, @0xDBFB7 on Twitter, or irc.0xDBFB7.com:6667 \#covid.}}
\date{May 2020}
\author{Based primarily on exceptional work by }

@@

Though the original research used Inf. A, our testing was only performed with a surrogate bacteriophage. All experiments must be repeated with SARS-NCoV-2.

Additionally, other mechanisms may produce superior results and should be evaluated in the same context. Recent data on far UVC [Buonanno 2017] indicate safety.

@@

@@

Because this is an ostensibly novel and niche mechanism, it may behoove us to breifly review the biological basis for the safety limits set by standards organizations. \footnote{The idea of biological microwave resonances appears to have originated in [Frohlich 1968, 1980].}

@@

The IEEE standard which the flagship paper [Yang 2015] references [IEEE C95.1-2005] was recently overhauled [IEEE C95.1-2019] with new metrics.

@@

There are some 20,000 papers on the topic of RF safety, with solid in vitro, in vivo and epidemiological evidence of safety in the 2.4 GHz and 5.8 GHz bands.

However, "There are limited experimental human data upon which to set limits on exposures above 6 GHz" [Chan 2019]. Only 2\% of the above papers are on frequencies $>6$ GHz [Vijayalaxmi 2018].

In addition, what data remains in this category is of varying quality.

As has been starkly demonstrated in the recent hydroxychloroquine contradictions [Gautret 2020] [Geleris 2020], biological research must be essentially perfect to have any meaning at all.

[Vijayalaxmi 2018] is a remarkable meta-analysis of in vitro data. They synthesize a quality score, rating bindedness, sham controls, dosimetry, and sample size. Quality was inversely, monotonically related to the effect size. An example of the subtle effects that can call into question superfically reasonable results are [

They additionally find a publication bias of incredible magnitude in the positive (harm) direction in this field, further degrading the usefulness of statistical analyeses.

An example of a relevant study that they grade as "quality 1" [Karaca 2011]. This study's conclusion is that However, in figure 5, one can see that only 1 out of 11 genes tested had a statistically significant difference in expression. Given the N=6 cultures,

[Adair 2002] demonstrate theoretically that acoustic resonances are, in general, highly unlikely in biological solvents, primarily focusing on DNA. All resonance modes are strongly overdamped by the surrounding solvent, and no amplitude amplification can take place.

This analysis would seem to disagree with the viral resonance apparently conclusively demonstrated by [Yang 2015] - the PCR result seems especially inarguable. This could be explained by the large charge on the virus compared to the molecules analyzed in Adair. [Yang 2015] find a charge of order $10^7 e$, whereas Adair analyzes a very different regime, "assume coupling to the field through single charges q = e at each end".

[Liu 2009] cite [Edwards, Swicord 1984], saying: "It was proposed that a hydration layer surrounding DNA molecules could lower the viscoelastic transition
frequency, raise the quality factor of confined acoustic vibrations, and result in a microwave resonant absorption". But [Adair 2002]'s [Foster 1987] seem to quite conclusively put the kibosh on that idea, demonstrating no resonances in DNA with 20x higher precision and showing previous results to be an artifact of the measurement equipment.

The supplemental material to [Vijayalaxmi 2018] has a table of studies on this topic. High-quality research in-vitro concurs with.

@@

9.4 GHz pulsed / in vivo, mouse, $0.1-10 \text{mW}/\text{m}^2$ over the whole body for 2 weeks. N=16.

@@

More modern research [Juutilainen 2011], [Chemeris], [Vijayalaxmi 2006],  does not tend to agree with these concerns. As cited by [ICNIRP 2020]'s excellent literature review, expose lymphocyte blood cells to 8 ns pulsed 8.2 GHz radiation for 2 hours at an average power of $10 \text{mW/cm}^2$. They find no change in any of the parameters measured, including various chromosomal parameters. They further discuss previous results.

In the frequency range of interest, there seems to be good-quality in-vitro evidence using lymphocytes of lack of harm. We are not aware of any high-quality in vivo or epidemeological data from which useful conclusions can be drawn.

We must defer to health experts for interpretation of these data.

@@

@@

[Hand 1982] $E_{mag}=1/e$ (13.5\% power density) skin\footnote{electromagnetic skin, not tissue skin}\footnote{well, both, I suppose.} depth is approximately:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
F=10 GHz          & Dry tissue & Wet tissue \\ \hline
Penetration depth & 30 mm      & 5 mm \\ \hline
\end{tabular}
\end{center}

@@

However, destruction of lung tissue appears to be the primary cause of death via SARS [Nicholls 2006].

A minimally-invasive bronchoscopic technique may be effective when mounted on a modified brochoscope and periodically inserted into the main bronchus, a la [Yuan 2019].

The bronchi are less than 2 mm thick [Theriault 2018] and the lungs themselves are only on the order of 
7 mm thick [Chekan 2016].

Further study is required to validate this method.

@@

For ease of design and simulation, a device with Touchstone S-parameters and SPICE files is greatly preferable.

As a result of our gross incompetence, the oscillator was designed via an inane, roundabout, and fiendishly tedious manner, and our description of this technique only contributes to the field by being suitable for a dartboard; our analysis can only hold water when combined with paper mache; and a lesser invertebrate in posession of a copy of Microwave Oscillator Design would have accomplished the task faster.

Designing an oscillator of this type in one step with a few kindergarten equations appears to be well within the reach of modern network analytical techniques. Genetic algorithms are quite well matched to this problem, and many commercial software packages are  [computer microwave design book].

In addition, the design of wideband feedback-loop VCOs is a relatively well-explored field, and many reference designs exist.

The key sticking point - which we missed - seems to be that at these scales, using microstrip design techniques, the parasitics of any possible filter structure are so large that the small change in inductance that varactors can provide cannot tune the circuit by any meaningful amount.

"Using the lead inductances of the bipolar transistor and varactors provides the required value of the base inductance".

Indeed, [Tsuru 2008]'s "tuned circuit" in Fig 8 is, in fact, just the varactors, plus an almost invisible high-impedance line.

We began experimentally by building a scaled version of [Mueller 2008]'s active antenna in microstrip, using a CEL mosfet.

This immediately oscillated perfectly.

Several analytical filter design methods were

%
\paragraph{\textbf{The feedback loop oscillator}}\

@@

With 0.79 mm FR4 substrate and 0.2 mm (8 mil) wide traces, the maximum impedance achievable was about 115 ohms, which did not appear to be sufficient as an RF choke. Use of defected grounds can increase inductance, but this was not evaluated.

If suitably high-impedance traces are not available, a common technique is to use a quarter-wavelength line (approximately 6 mm long with the above parameters at 8 GHz) terminated with a stub to produce a virtual short or open circuit [Seo 2007].

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\Large Molecular dynamics simulation}\\
\begin{multicols}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

aa

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\Large Optical centrifuge polarization chirp}\\
\begin{multicols}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

aa

@@

Professor Shahramian's excellent video blog {\it The Signal Path} inspired this work. Alex Dainis' videos on sterile technique were also useful.

@@

They say: "Based on the IEEE Microwave Safety Standard, the spatial averaged value of the power density in air in open public space shall not exceed the equivalent power density of $100(f/3)^{1/5} \text{ W/m}^2$ at frequencies between 3 and 96GHz ($f$ is in GHz). This corresponds to $115 \text{ W}/\text{m}^2$ at 6 GHz". The 115 W @ 6 GHz value correctly corresponds to this equation with a coefficient of 100.

@@

We don't think this changes the substance of [Yang 2015]'s paper, if it is an error; the relaxed localized exposure is probably the correct threshold to use, and it just means the mechanism is more useful for healthcare workers who acknowledge the risks.

@@

Simulating the oscillator with an AC sim using the manufacturer's S-parameters and QUCS' microstrip approximations [qucs/optimize\_filter\_1] intially appeared to yield good agreement with experiment. Peaks in the feedback voltage simulation corresponded approximately with peaks in the observed spectrum. [figures from LO prototype N].

@@

A feedback loop filter design with apparently ideal phase and frequency properties was designed using this process; but the behavior of this oscillator did not correspond to reality.

@@

Because we lacked a spectrum analyzer that could monitor the higher poles of the filter until the bootstrap LO was designed, we had to rely on in-silico analyeses.

@@

As of 0.0.20, QUCS' microstrip models are not yet compatible with transient simulations; and some improved filter designs required simulating coupling between more than two microstrips, which QUCS did not yet support natively. Coupling can be emulated by introducing discrete coupling capacitors

Transient simulations with ngspice matched experiment far more closely.

@@

There are many ways of linking SPICE and FDTD.

@@

Phase shift can be introduced either via a length of microstrip, or a high-pass/low-pass filter [Microwave101]. Adding a fixed microstrip line restricts the tuning range, however, and the filter inevitably affects the frequency response.

@@

It may be helpful to think of simulations very similarly to IRL experiments.

A simulation is just a universe in a bottle that you can examine more closely; as with real-world, you can learn much from observation, but

But this need for documentation conflicts with the rapid, iterative cycle necessary for productivity.

This is obvious to all compentent, but when rapidly iteratively testing with simulations, it may be helpful to automatically save a package with images of all the components (schematics, graphs, input files) of each distinct test.

For comparison to experiment, having a webcam take an image of the assembled board is also helpful.

Version control alone isn't quite enough. Just having a simulation setup file somewhere in the commit history isn't "discoverable" - that is, you must be able to see what the input and output was without re-running the simulation.

@@

Software such as Sumatra, Sacred, recipy, and others. In our case, we used eLabFTW's elabapy bindings.

@@

For instance, this was done with the PDB reader in the OpenMM wrapper and JAMES.

There's also a very neat thing that seems to be common in computational biology, but which doesn't seem popular in other fields. After an algorithm or software tool is written (and the source published seperately), a simple CGI frontend is written around it and hosted on the university's servers. eLNemo and the CHARMMing web interface are advanced examples of this. This way, anyone with an input file can get results without futzing around with installation.

The computational biologists have beat us at our own game.

@@

We have always encountered wasting extreme amounts of time on subtle assembly mistakes in hardware prototypes.

In one example, many hours were spent because the enamel insulation on a bodge wire had not burned off completely in the solder joint, leading to a high-impedance connection.

The same has occurred in previous projects; in one example, many days were spent debugging software to fix an apparently slow hardware interrupt, which ended up being the result of a poor solder paste stencil leading to a hidden high-impedance connection to a leadless package.

Many failures are perhaps the result of carelessness in modification and a lack of inspection; but others would only have been found by a 100\% electrical test.

If flying-probe or bed-of-nails tests can be made sufficiently rapid and closely coupled with the existing toolchain,

In {\it 2001: A space odyssey}, an automated system is shown guiding the troubleshooting of an assembly, apparently generating a fault tree of all the

@@

We present this only as a cautionary tale in the hopes that someday I will listen.
